### 2014-06-12 Paper1/paper1.tex
Shouldn't fig5 be deleted?

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test it on human reaching data. Popular feedforward control models     approaches can be inverted in a general approach, and we show here the strengths, limitations, and underlying assumptions. We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen in human targeted reaching. Further, most of the sensitivity resulted from uncertainty with regard to the joint stiffness. We also tested the robustness of this approach on human point-to-point reaching movements disturbed by large force pulses. As expected, the intended trajectory showed no change from undisturbed reaching for 150 ms after the disturbance onset. Beyond this point, intent change, arm stiffening, or both usually lead to deviation from undisturbed reaching after an average of 300 milliseconds after the onset of the disturbance, even when the intention was near the target.  Unlike the well-known $\lambda$ models, this intent determined by considering a both feedforward (predictive) and feedback controls. Knowing the intent along with the action should allow real-time determination of error, making it possible to enhance error, and also directly inspect a neurally-impaired patient's intent for deficits. This approach may also be scaled for determining intent in complex systems of team sports, crowds, or swarms.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
In our everyday struggles with bumpy roads and bustling crowds we often make movements in ways other than our plan. Here we identify a computational modeling method that measures actions and allows us to determine what the person meant to to do. We first demonstrate this method's good performance even when there is uncertainty in the parameters of the model. Next, we evaluate this on human reaching motions to targets, and show that while the arm is bumped by a disturbance, the intention remains straight. Approximately 300 milliseconds later, the intention changes in a corrective action that counteracts the disturbance. This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from molecules to crowds of people. It is also a new tool for understanding and treating impaired movement plans in neural injury.

\section*{Introduction}
As machines interact more and more with humans, there is a growing need for a rigorous way to understand what people are trying to do, even if they become disturbed, distracted, or influenced by pathology. One can assume that a person who is familiar with a particular task will try to do the same thing each time, provided all other conditions are equal. This presumes that what people do is on average what they intended to do and that their goal remains unchanged. However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention. Challenges arise in human-machine interactions because intention and changes in intention cannot be easily discerned. Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.

@@

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point. Supporters of the so-called $\lambda$ model\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement. Upon inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. We propose a revised method---intent extraction---that deduces intended trajectory, even in the face of disturbances. This method describes a class of filters that map a disturbance, process, actuator, and realized trajectory to an intended trajectory. In human reaching, the process (a dynamic model of the arm), the actuator (viscoelastic muscles), and the realized trajectory are all readily accessible via measurement and modeling. The underlying framework should be extensible to whole-body movement or even more abstract, psychological processes.

Collectively, the general problem of being able to best extract someone's intent is critical to many fields, and hinges on our understanding of the underlying mechanisms through which control is accomplished. We have excellent models of the mechanical properties of arm and muscle because the physics that describe their gross behavior is so well-understood and the need to determine particular physical parameter values has long been clear. This wealth of previous work provides strong footing for our technique, which in turn should yield valuable information not previously available by other means. In fields where the process and actuators are not yet well-understood, this approach may help to structure the search for them.

Here we develop and analyze a method to extract intent from disturbed movement, which will allow us to observe if and how intent changes following disturbance. We employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction. We then explore sensitivity and failure modes using standard model analysis techniques. We also demonstrate the accuracy and utility of this on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances. We hypothesize that this combination allows intent extraction with less uncertainty than the uncertainty naturally inherent in point-to-point reaching. We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for a providing a better understanding of what a person means to do when they move and are disturbed.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Methods}
The sections below describe the theoretical method, and then present our evaluations. First, we considered an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process. Next we conducted an experiment on humans and the process was then evaluated on human subject data.

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation. Additionally, linear feedback rejects position and velocity error. This model is chosen to be very simple to help illuminate the approach, but it is limited in how it captures all of the complexity of motor control. Nevertheless later we will discuss how this might be used on more complex models without loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
Where $M$ is the mass matrix, $q$ is the joint angles, $G$ contains both Coriolis and centripetal effects, and $E$ is externally-applied torque that the system might interact with. Dots ($\dot{q}, \ddot{q}$) are used to denote time derivatives of the joint angles. The motion behavior changes with the addition of feedforward and/or feedback controllers:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
\end{equation}
Where a hat ( $\hat{ }$ ) over term indicates the nervous system's best estimate of the planar dynamics, also known as the internal model. This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm. If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero. $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$. This $q_d$ signifies the unknown desired trajectory that we seek to discover.

For typical dynamic simulations, in order to determine the trajectory of the system (i.e., the forward dynamics problem), this second-order differential equation requires standard numerical integration to determine the solution to the initial value problem in time. This first involves algebraic manipulation to solve for $\ddot{q}$, and then integration to determine the state trajectory. The intent extraction approach modifies this method by instead solving for $\ddot{q}_d$:
\begin{equation}
\ddot{q}_d=\hat{M}(q_d)^{-1}\left\{M(q)\ddot{q}+G(q,\dot{q})+E-[\hat{G}(q_d,\dot{q}_d)+\hat{E}+K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})]\right\}
\end{equation}  
$\ddot{q}_d$ can then be integrated to determine the intended state trajectory $q_d(t)$. Provided the model of plant and controller are precise, the initial conditions are available and accurate, the mass matrix estimate $\hat{M}$ is invertible, and the constants are accurate and do not vary, and externally-applied force can be accurately measured, then the system yields an accurate estimate of the intent.

It is important to distinguish our intended trajectory, $x_d$, from another variable often considered, the equilibrium point of the muscle actuators. Importantly, this has been claimed to be the intended movement. To consider this, we also generalize our approach to any system with a moving intention that governs the control law. This introduces two additional layers of complexity by separating the plant (arm) into its process (dynamics) and actuator components (muscles) and considering the actuator's equilibrium, $x_e$ (sometimes represented as $\lambda$). The feedforward component must use and share this actuator and hence use the actuator's equilibrium. Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some space, $x$:
\begin{equation} \label{eq:generalPlant}
\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
\end{equation}
The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ that $x$ will track a desired path $x_d$.    
\begin{equation}
\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m(x_d-x_e)^{(m)}}^\text{Actuator}=0
\end{equation}
We solve this for $x_e$ and substituting into \eqref{eq:generalPlant} gives:
\begin{equation}
\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+E=\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m (x_d-x)^{(m)}}^\text{Feedback (Actuator)}
\end{equation}

Note $x_e$ vanishes, recovering our familiar model. While a proper choice of $x_e$ is critical to control, our approach does not require a model relating equilibrium and intended trajectory. Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero. This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}. In summary, if the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.

While we treat both the process and actuator as dynamic models in our derivation, both the derivation and the technique are agnostic to the process model. Instead of a dynamic model, it could be a lookup table. For instance, a light switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism. In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance. As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined. This allows determination of intention in many situations of interest, even where the process is irreducibly complex.

\subsection*{Generation of synthetic data using simple dynamic simulation}
We choose one prevailing arm model, that of Burdet et al. \cite{burdet2006stability}, along with the anatomical mass distribution and geometry parameters of Winter \cite{winter2009biomechanics} to create a model for estimating intent. The model of Burdet et al \cite{burdet2006stability} was adapted to reflect measured physical parameters of the subject who provided the reach. A scalar multiple of feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation}. Segment lengths were measured directly while gross body mass was self-reported. These values were converted to segment masses, centers of gravity, and moments of inertia using anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics}. The muscle torque used to calculate stiffness was approximated by looking backward in time a small amount (500 microseconds). Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration.

Simulated reaches were 15 centimeters long, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then proceeding right.  We used the following types of perturbing forces for each combination of distance and direction: 
pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion. The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.

\subsection*{Error Metric}
Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15centimeters reach.  Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis. Mean and maximum signed error relative to the direction of disturbance, also called perpendicular deviation, were used to measure reaching accuracy in order to elucidate direction of any corrections.

\subsection*{Indices of variance-based sensitivity}
Sobol-distributed matrices were generated using MATLAB's sobolset() function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density. Distributions were clamped to the range of $\pm3$ standard deviations in order to avoid support for implausible parameter values. Combinations and calculations were made as prescribed by Saltelli et al. \cite{saltelli2010variance} to arrive at direct and total sensitivity indices.

@@

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any handle friction or mass. The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback was performed at 60 Hz using OpenGL. Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.

@@

\subsection*{Model Robustness}
Sensitivity analysis fortunately revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.  Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (Figure 1) were extracted in the presence of 220,000 variations upon the parameters (Table 1) according to the methods of Saltelli et al. \cite{saltelli2010variance}. Expected variance due to parameter uncertainty was $0.86  \mu m^2$ for pulse forces and $1.5 \mu m^2$ for filtered white forces. Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mu m^2$. These variances all fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter). In other words, while some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input.

\subsection*{Parameter Sensitivity}
Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial. Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all parameters were accurate,  and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty. Note that this sensitivity analysis does not provide information on whether models or parameter estimates are accurate, only how sensitive the model would be if they were inaccurate.

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3). In both pulse timing conditions, intent remained straight line towards the target, supporting the approach.  When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred late in reaching, intent continued onto the target (Figure 3 lower plots). We observed no significant change of intent during force disturbance until 315 milliseconds after the onset of disturbance. Despite statistical insignificance, intention showed some systematic deflection toward the direction of disturbance beginning around 275 milliseconds after the onset of force disturbance. While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to first over-correct and is altered for almost 900 milliseconds (Figure 4 shaded areas).

As this is the first sound method for looking at intent following disturbance, we have only the undisturbed intention for comparison. If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands. This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay. We test this in Figure 4.

Following disturbance, we compare deviation in the hand and intended paths against the maximum deviation of undisturbed movements in order to determine at what point in time changes in hand path or intent can no longer be explained by normal variation and inaccuracy. Subjects' force-disturbed hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement beginning within at most 135 milliseconds after the onset of disturbance. By contrast, intent did not change significantly for at least 300 milliseconds after the onset of disturbance (Table 2). This passes the test described above. While there does appear to be some systematic failure of the intent model to completely reject the disturbance, this is not significant. An overcorrection beginning at approximately 200 milliseconds appears to account for change in intention in all subjects.

>>>>

% Please keep the abstract between 250 and 300 words
\section*{Abstract}
We derive a novel modeling approach for determining movement intent from a force-disturbed planar arm trajectory, examine its parameter sensitivity, and test the approach using human reaching data. We first show how to mathematically invert typical feedforward control approaches, and then examine how this reveals the strengths, limitations, and underlying assumptions. We found that sensitivity to parameter inaccuracies was on the order of microns and hence much lower than the inaccuracies seen when humans reach to visual targets. Further, most of the sensitivity resulted from uncertainty with regard to the joint stiffness. We also tested the robustness of this approach on human point-to-point reaching movements disturbed by force pulses. As expected, the intended trajectory showed no change from undisturbed reaching for 300 ms after the disturbance onset. Beyond this point, we detected change in intent even when the intention was near the target.  Unlike well-known $\lambda$ models, this intent was determined by considering both feedforward (predictive) and feedback controls. Knowing intent alongside action allows real-time determination and manipulation of error, even without a specified task, which makes it possible to directly inspect a neurally-impaired patient's intent for deficits. This approach may also be scaled for determining intent during human-machine interaction and in complex systems like team sports, crowds, or swarms.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}
In our everyday struggles with bumpy roads and bustling crowds we often make movements that do not match what we had planned. Here we identify a computational modeling method that measures actions and allows us to determine what the person meant to to do. We first demonstrate this method's good performance even when there is uncertainty in the parameters of the model. Next, we evaluate this on human reaching motions to targets, and show that while the arm is bumped by a disturbance, the intention remains straight. Approximately 300 milliseconds later, the intention changes in a corrective action that counteracts the disturbance. This type of computation can be broadly useful in canceling the effects of disturbances such as turbulence while typing or piloting, and can estimate intent in many other dynamic systems from molecules to crowds of people. It is also a new tool for understanding and treating impaired movement plans in neural injury.

\section*{Introduction}
There are many situations, such as disturbances, distractions, and pathologies, that interfere with execution of an intended movement. One can assume that a person who is familiar with a particular task will try to do the same thing each time, provided all other conditions are equal. This presumes that what people do is on average what they intended to do and that their goal remains unchanged. However, if there is a disturbance, it is unclear what happens to a person's intentions because action is decoupled from intention. Challenges arise in human-machine interactions because intention and changes in intention cannot be easily discerned. Hence a comprehensive method for understanding what someone means to do (rather than what they actually do or claim to have wanted to do) as they interact within uncertain environments will prove useful in any area where a person might be thwarted from their intent such as national security, operation of machinery, athletic and music performance, rehabilitation, human augmentation, and artificial intelligence.

@@

Attempts to deduce motor intent in the past have focused on the assumed spring-like properties of human muscles. Springs produce a force according to their impedance and stretch. By measuring force, impedance, and position, Gomi and Kawato\cite{gomi1997human} were able to deduce stretch and thereby infer the muscle's equilibrium point. Supporters of the "$\lambda$ model"\cite{feldman1995origin} hypothesized that this equilibrium point represented the intent of a movement. Upon Gomi and Kawato's inspection of the equilibrium point as it evolved in time, it was clear that it was highly complex and often not anatomically realizable. Therefore, it could not well-represent the intent of a simple reaching movement. We propose a revised method---intent extraction---that deduces intended trajectory, even in the face of disturbances. This method describes a class of filters that map a disturbance, process, actuator, and realized trajectory to an intended trajectory. In human reaching, the process (a dynamic model of the arm), the actuator (viscoelastic muscles), and the realized trajectory are all readily accessible via measurement and modeling. The mathematics underlying intent determination are sufficiently general to be extensible to whole-body movement or even more abstract, psychological processes.

Here we develop and analyze a method to extract intent from disturbed movement. This method then allows us to observe if and how intent changes following disturbance. We employ bidirectional modeling: we convert intent into movement using a forward model so that we can examine the quality and uncertainty of converting movement to intent by intent extraction. We then explore sensitivity and failure modes using standard model analysis techniques. We also demonstrate the accuracy and utility of this on a preliminary dataset of healthy reaching movements disrupted by occasional, unpredictable disturbances. We hypothesize that this combination allows intent extraction with less uncertainty than the uncertainty naturally inherent in point-to-point reaching. We show that certain factors and measurements are more sensitive than others, and demonstrate the viability of this method for providing a better understanding of what a person means to do when they move and are disturbed.

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Methods}
The sections below describe the theoretical method and the means by which we evaluate it. First, we considered an idealized system using synthetic data that demonstrated the concept and provided an understanding of the computational process. Next we conducted an experiment on humans to evaluate the process using human subject data.

\subsection*{Intent Extraction: Deducing the Desired Trajectory}
We first demonstrate the \textit{intent extraction} approach in human motor control, but we later show that the process is applicable to any controlled process with an invertible highest order plant term. The process begins by presuming a model of the controller. Here, we choose the well-known motion control structure of Shadmehr and Mussa-Ivaldi\cite{shadmehr1994adaptive} where the feedforward aspect of the controller perfectly predicts the plant and linearizes the system through cancellation. Additionally, linear feedback rejects position and velocity error. This model was chosen to help illuminate the approach, but its feedback component lacks the complexity and accuracy of more recent approaches. Nevertheless later we will discuss how this approach can be used on more complex models without loss of generality. The equation governing the passive planar dynamics (plant) of musculoskeletal structure is of the form:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=0
\end{equation}
Where $M$ is the mass matrix, $q$ is the joint angles, $\dot{q}$ is joint angular velocity, $\ddot{q}$ is joint angular acceleration, $G$ contains both Coriolis and centripetal effects, and $E$ is any externally-applied torque. The motion behavior changes with the addition of feedforward and/or feedback controllers:
\begin{equation}
\underbrace{\overbrace{M(q)\ddot{q}}^{\text{Inertia}}+\overbrace{G(q,\dot{q})}^{\text{Coriolis, Centripal}}}_{\text{Plant}}+E=\underbrace{\overbrace{\hat{M}(q_d)\ddot{q}_d}^{\text{Inertia}}+\overbrace{\hat{G}(q_d,\dot{q}_d)}^{\text{Coriolis, Centripal}}+\hat{E}}_{\text{Feedforward Controller}}+\underbrace{K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})}_{\text{Impendance, Feedback Controller}}
\end{equation}
Some terms have hats over them ($\hat{M}$, $\hat{G}$, or $\hat{E}$) to indicate that they represent the nervous system's best estimate of the forces and dynamics it will encounter, which is also known as an internal model \cite{kawato1999internal}. This portion of the system serves as an inverse-dynamics feedforward controller that cancels out the dynamics of the arm in the torque balance. If the nervous system has sufficient experience and is expecting $E$, it is included as part of the internal model, $\hat{E}$, otherwise $\hat{E}$ is set to zero. $K_p$ and $K_d$ are the lumped impedance and feedback terms that employ a moving state equilibrium to accomplish the desired trajectory, $q_d$. This $q_d$ signifies the unknown desired trajectory that we seek to discover.

For typical dynamic simulations, in order to determine the trajectory of the system (i.e., the forward dynamics problem), this second-order differential equation is typically solved by numerical integration to determine the solution to the initial value problem in time. This first involves algebraic manipulation to solve for $\ddot{q}$, and then integration to determine the state trajectory. Intent determination takes the novel approach of instead solving for $\ddot{q}_d$:
\begin{equation}
\ddot{q}_d=\hat{M}(q_d)^{-1}\left\{M(q)\ddot{q}+G(q,\dot{q})+E-[\hat{G}(q_d,\dot{q}_d)+\hat{E}+K_p(q_d-q)+K_d(\dot{q}_d-\dot{q})]\right\}
\end{equation}  
$\ddot{q}_d$ can then be integrated to determine the intended state trajectory $q_d(t)$. This relies on many assumptions: The model of plant and controller must be accurate and precise. The initial conditions must be available and accurate. The mass matrix estimate, $\hat{M}$, must be invertible. Externally-applied force must be precisely and accurately measured. If all of these conditions are met, then the system yields an accurate estimate of the intent.

It is important to distinguish our intended trajectory, $q_d$, from another variable often considered, the equilibrium point of the muscle actuators. Importantly, this has been claimed to be the intended movement. To consider this, we also generalize our approach to any system with a moving intention that governs the control law. This introduces two additional layers of complexity by separating the plant (arm) into its process (dynamics) and actuator components (muscles) and considering the actuator's equilibrium, $q_e$ (sometimes represented as $\lambda$). The feedforward component must use and share this actuator and hence use the actuator's equilibrium. Without loss of generality, we represent the process $P$ and actuator $A$ as linear (or linearizable) operations in some space, $x$:
\begin{equation} \label{eq:generalPlant}
\underbrace{\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+\overbrace{\sum_{m=1}^M A_m(x-x_e)^{(m)}}^\text{Actuator}}_\text{Plant}+E=0
\end{equation}
The internal model of the process ($\hat{P}$ and $\hat{E}$) predicts system actions, external disturbances, and impedance responses in order to determine $x_e$ that $x$ will track a desired path $x_d$.    
\begin{equation}
\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m(x_d-x_e)^{(m)}}^\text{Actuator}=0
\end{equation}
We solve this for $x_e$ and substituting into \eqref{eq:generalPlant} gives:
\begin{equation}
\overbrace{\sum_{n=1}^N P_nx^{(n)}}^\text{Process}+E=\overbrace{\sum_{n=1}^N \hat{P}_n x^{(n)}_d+\hat{E}}^\text{Internal Model}+\overbrace{\sum_{m=1}^M A_m (x_d-x)^{(m)}}^\text{Feedback (Actuator)}
\end{equation}

Note $x_e$ vanishes, recovering our familiar model. While a proper choice of $x_e$ is perhaps essential for control, our approach does not require a model relating equilibrium and intended trajectory. Note also that the actuator's equilibrium $x_e$ is not the process's equilibrium unless all derivatives of $x_d$ are zero. This upholds Gomi and Kawato's finding that the arm muscles' equilibrium does not represent reaching intent\cite{gomi1997human}. In summary, if the highest order coefficient of the dynamic model of the process $\hat{P}_N$ can be inverted and the impedances can be modeled, it is possible to solve for $x_d^{(N)}$ and integrate for $x_d$ revealing the intended trajectory.

While we treat both the process and actuator as dynamic models in our derivation, it is important to note that both the derivation and the technique are agnostic to the process model. Instead of a dynamic model, it could be a lookup table. For instance, a light switch is well-modeled as a lookup table or threshold without any consideration of its underlying mechanism. In this case, our technique could be used to determine a person's intention to flip the switch and how it changes in the face of disturbance. As long as some bidirectional relationship (even statistical) exists between state and outcome, intended outcome can be determined. This allows determination of intention in many situations of interest, even where the process is irreducibly complex.

\subsection*{Generation of synthetic data using simple dynamic simulation}
While the derivation above is general, testing its application to human reaching requires choices for plant and actuator models. Appropriate plant modeling is well-understand: measured arm segment lengths and self-reported body mass are converted into the plant's inertial, centripetal, and Coriolis terms. Anatomical landmarks and values from Dempster \cite{dempster1955space} and Winter \cite{winter2009biomechanics} relate body mass to limb mass, limb length to limb center of mass, and limb mass and length to moment of inertia. For the actuator, we choose the model of Burdet et al. \cite{burdet2006stability}. A scalar multiple of their modeled feedback torque, determined by linear regression, was used to account for any task-dependent differences in impedance \cite{franklin2003adaptation} while leaving the shape and structure of the impedance unchanged. As a consequence of choosing this model, muscle torque must be used to calculate stiffness even though it arises in part from stiffness. Like Burdet et al. \cite{burdet2006stability}, we approximated muscle torque by looking backward in time a small amount (500 microseconds). This approximation is not necessary during intent extraction, which leads to a small (sub-micrometer mean unsigned error) discrepancy. Desired trajectory in time was simulated as a minimum jerk, 5th order polynomial of duration 700 milliseconds starting and ending with zero velocity and acceleration.

\subsection*{Experimental Design}
Simulated reaches were 15 centimeters long, beginning 38 centimeters out from and 5.7 centimeters left of the right shoulder then proceeding right.  We used the following types of perturbing forces for each combination of distance and direction: 
pulse forces applied in one of the the two directions perpendicular to the direction of movement began when the subject had moved either 10\% or 50\% of the distance to the target and lasted for 150ms.
Noise forces began once the subject moved 3 millimeters, and lasted for the duration of the motion. The forces were drawn from a white noise generator at 1000 Hz with flat power spectral density of 1N, and then passed through a 4th order low-pass Butterworth filter with cutoff $10\pi$ rad/s.

\subsection*{Error Metric}
Our error metric used deviation in position, where unsigned error was calculated as the mean magnitude of deviation from the nominal trajectory across each of 150 bins measuring 1 millimeter in width and spaced evenly throughout the 15 centimeter reach.  Mean unsigned error (MUE) was also used to summarize the overall error in each movement for sensitivity analysis. Mean and maximum signed error relative to the direction of disturbance, also called perpendicular deviation, were used to measure reaching accuracy in order to elucidate direction of any corrections.

\subsection*{Indices of variance-based sensitivity}
Sobol-distributed matrices were generated using MATLAB's sobolset() function and converted to parameter distributions (see Table~\ref{tab:parameters}) using inverse cumulative probability density. Distributions were clamped to the range of $\pm3$ standard deviations in order to limit parameters to a realistic range. Combinations and calculations were made as prescribed by Saltelli et al. \cite{saltelli2010variance} to arrive at direct and total sensitivity indices.

@@

\subsection*{Apparatus}
A planar manipulandum (described in Patton and Mussa-Ivaldi \cite{patton2004robot}) was programmed to compensate and minimize any friction or mass. The MATLAB XPC-TARGET package \cite{MATLAB:2008} was used to render this force environment at 1000 Hz and data were collected at 1000 Hz.  Visual feedback was performed at 60 Hz using OpenGL. Closed-loop data transmission time (position measurement to completed rendering to recognition of rendering by the position measurement system) was less than 8 milliseconds, ensuring a visual delay less than one 60 Hz frame.

@@

\subsection*{Model Robustness}
While some inversion processes might be highly sensitive to model uncertainty, our process for recovering intent from action is much more precise than its input. Sensitivity analysis revealed that variance in intended trajectories due to estimated uncertainty in model parameters is lower than the natural variation in undisturbed motions.  Simulated point-to-point reaches disturbed by either filtered white noise forces or a pulse force perpendicular to the direction of the reach (Figure 1) were extracted in the presence of 220,000 variations upon the parameters (Table 1) according to the methods of Saltelli et al. \cite{saltelli2010variance}. Expected variance due to parameter uncertainty was $0.86  \mu m^2$ for pulse forces and $1.5 \mu m^2$ for filtered white forces. Variance in recorded undisturbed point-to-point reaching under the same time and reach distance conditions was $2.93 \mu m^2$. These variances all fall beneath the instantaneous accuracy of the robot recording the motion (tenths of a millimeter).

\subsection*{Parameter Sensitivity}
Simulated error due to direct parameter uncertainty (Figure 2) reached the order of millimeters and revealed particular sensitivity to misestimation of stiffness and changes in shoulder position from trial to trial. Direct sensitivity, which reveals the proportion of error that could be removed by correcting inaccuracy of a parameter, was an order of magnitude lower than total sensitivity, which reveals how much error would remain if all other parameters were accurate,  and both implicated stiffness estimation inaccuracy as a primary cause of extraction uncertainty. Note that this sensitivity analysis does not provide information on whether models or parameter estimates are accurate, only how sensitive the model would be if they were inaccurate.

\subsection*{Human Intent Trajectories}
Extraction of intent from human point-to-point reaching revealed straight-line movement that persisted for hundreds of milliseconds after the onset of disturbing forces (Figure 3). In both pulse timing conditions, the intended trajectory remained straight towards the target, supporting the approach.  When disturbances occurred early in reaching, intent showed some signs of correction mid-reach. When disturbances occurred late in reaching, intent continued onto the target but then corrected anyways (Figure 3 lower plots). We observed no significant change of intent during force disturbance until 300 milliseconds after the onset of disturbance. Intention showed some systematic deflection toward the direction of disturbance peaking around 150 milliseconds after the onset of force disturbance, but this effect did not leave the range of error found in undisturbed movement. While the hand returns within 600 milliseconds to undisturbed levels of error, intention appears to first over-correct and is altered for as much as one second (Figure 4 shaded areas).

We argue that the approach here represents the first sound method for looking at intent following disturbance, so we have only the undisturbed intention for comparison. If intent can change following disturbance, it can only do so after some period of time due to delay in processing sensory information and producing new motor commands. This represents a way to falsify the method: our determined intent should not diverge from the undisturbed intent of similar motions within this window of delay. We test this in Figure 4.

Following disturbance, we compare deviation in the hand and intended paths against the maximum deviation of undisturbed movements in order to determine at what point in time changes in hand path or intent can no longer be explained by normal variation and inaccuracy. Subjects' force-disturbed hand paths could no longer be explained ($p=0.05$) by deviation in undisturbed movement beginning within at most 135 milliseconds after the onset of disturbance. By contrast, intent did not change significantly for at least 300 milliseconds after the onset of disturbance (Table 2). This passes the test described above. While there does appear to be some systematic failure of the intent model to completely reject the disturbance, this residual disturbance does not exceed the statistical threshold for significance. An overcorrection beginning at approximately 200 milliseconds appears to account for change in intention in all subjects.
